Testing atlantisom: end to end truth for California Current

### 2019-05-22 Atlantisom2SSworkflowtest.Rmd
update CCA workflow with correction, biol sampling

# should return a perfectly scaled survey 
effic1 <- data.frame(species=funct.group.names,
                     efficiency=rep(1.0,length(funct.group.names)))

# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
# BUT CHECK if newer Atlantis models can do age-specific outputs
selex1 <- data.frame(species=rep(funct.group.names, each=10),
                     agecl=rep(c(1:10),length(funct.group.names)),
                     selex=rep(1.0,length(funct.group.names)*10))

# should return all model areas
boxpars <- load_box(d.name, box.file)
boxall <- c(0:(boxpars$nbox - 1))

# generalized timesteps all models
runpar <- load_runprm(d.name, run.prm.file)
noutsteps <- runpar$tstop/runpar$outputstep
stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc

timeall <- c(0:noutsteps)

# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
# for ecosystem indicator work test all species, e.g.
survspp <- funct.group.names

# for length and age groups lets just do fish and sharks
# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)

survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
                           funct.groups$GroupType %in% c("FISH", "SHARK")]

surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

@@

@@

# first check that codebase is prior to corrected--to implement in atlantisom
#   read code version or date from log file
#   check against update date (Dec 12, 2015)

@@

# divide the numbers at age by (86400 * number_water_column_layers_in_the_box)

  catch <- catch.tmp %>%
    mutate(atoutput = atoutput * 86400 * numlayers) %>%

# replace truth$catch atoutput with correction

>>>>

@@

## Catch output correction--apply to all species

@@

# first check that codebase is prior to corrected--to implement in atlantisom
#   read code version or date from log file
#   check against update date (Dec 12, 2015)
# grep("Atlantis SVN Last Change Date", readLines(file.path(d.name, "log.txt")), value = TRUE)

# alternative: less overhead if the system does it? log file is huge
# logfile <- paste0(file.path(d.name, "log.txt"))
# codedate <- system(paste0("grep 'Atlantis SVN' ", logfile))
# compare date in codedate in an if statement--fix in function

@@

# divide the numbers at age by (86400 * number_water_column_layers_in_the_box)
# replace truth$catch atoutput with correction 
  catch <- catch.tmp %>%
    mutate(atoutput = atoutput / 86400 * numlayers) %>%
    select(species, agecl, polygon, time, atoutput)

  truth$catch <- catch

#save for later use, ar (better) resave truth file
#saveRDS(catch, file.path(d.name, paste0(scenario.name, "catchnums_corrected.rds")))
  save(truth,
      file = file.path(dir, paste0("output", scenario, "run_truth.RData")))
```

Did the correction work? Not sure what to compare this to. First put it back through `calc_biomass_age` even though this won't match either?

```{r correctcatage-tobio, eval=FALSE}

biol <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file)

# call within run_truth
catchbio <- calc_biomass_age(nums = catch,
    resn = truth$resn, structn = truth$structn, biolprm = biol)

#save for later use, even if wrong? takes a long time to generate
saveRDS(catchbio, file.path(d.name, paste0(scenario.name, "catchbio_ccatage.rds")))

It still doesn't match, similar to NOBA, but in the right ballpark now so the error is the median structn and resn called in the calc function above. This suggests that the catch at age in numbers is now correct, however.

```{r plotcorrected-bio}

catchbio <- readRDS(file.path(d.name, paste0(scenario.name, "catchbio_ccatage.rds")))

catchbio_corrected_agg <- aggregate(atoutput ~ species + time,
                                    data=catchbio, sum)

compare_catchB <-ggplot() +
  geom_line(data=catchbio_corrected_agg, aes(x=time,y=atoutput, color="catch atlantisom"), 
            alpha = 10/10) +
  geom_point(data=atCtxttidy, aes(x=Time/365,y=catchbio, color="txt output catch bio"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compare_catchB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
compare_catchB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
compare_catchB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
compare_catchB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")

## Fishery biological sampling

Now we can sample fishery catch for lengths, ages, weights:

Here we use `create_fishery_subset` on the numbers output of `run_truth` to create the survey census of age composition (for just our three species in this case). The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.

Because we don't want to wait 24 hours for this, we will look at only our three speces and the first 112 time steps.

```{r catchage-len-wt-3spp, echo=TRUE, warning=FALSE, message=FALSE}

# get survey nums with full (no) selectivity
catch_testNss <- create_fishery_subset(dat = truth$catch,
                                 time = c(0:111),
                                 species = spp.name,
                                 boxes = boxall)

# apply default sample fish as before to get numbers
catch_numssshigh <- sample_fish(catch_testNss, effNhigh)

# aggregate true resn per survey or fishery subset design
catch_aggresnss <- aggregateDensityData(dat = truth$resn,
                                 time = c(0:111),
                                 species = spp.name,
                                 boxes = boxall)

# aggregate true structn per survey or fishery subsetdesign
catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
                                 time = c(0:111),
                                 species = spp.name,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
catch_structnss <- sample_fish(catch_aggstructnss, effNhigh, sample = FALSE)

catch_resnss <-  sample_fish(catch_aggresnss, effNhigh, sample = FALSE)

This is true catch at age for sardine:

```{r catage1}

catchage <- catch_numssshigh %>%
  filter(species == "Pacific_sardine")

catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        catchage$species))

catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")

This is true catch at age(class) for hake:

```{r catage2}

catchage <- catch_numssshigh %>%
  filter(species == "Mesopel_M_Fish")

catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        catchage$species))

catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")

Hake will need `calc_stage2age` applied because it has 2 true ages per age class.

Length sample with user specified max length bin (200 cm):

```{r userset-maxlen, echo=TRUE}

catch_length_census_ss <- calc_age2length(structn = catch_structnss,
                                 resn = catch_resnss,
                                 nums = catch_numssshigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 maxbin = 200,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

We should get the upper end of anything with a 200cm max length bin.

Sardine catch lengths:

```{r vis-fishery-lf-test-1}

catchlen <- catch_length_census_ss$natlength %>%
  filter(species == "Pacific_sardine")

lfplot <- ggplot(catchlen, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        catchlen$species))

lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")

Hake catch lengths:

```{r vis-fishery-lf-test-2}

catchlen <- catch_length_census_ss$natlength %>%
  filter(species == "Mesopel_M_Fish")

lfplot <- ggplot(catchlen, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        catchlen$species))

lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")

Fishery weight at (st)age:

```{r vis-fishery-wtageclass-test}

wageplot <- ggplot(catch_length_census_ss$muweight, aes(agecl, atoutput)) +
  geom_point(aes(colour = time)) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  scale_x_discrete(limits=c(1:10)) +
  xlab("age class") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste(scenario.name))

wageplot + facet_wrap(c("species"), scales="free_y")

Change in wt at (st)age in the fishery over time for age classes using an annual mid-year snapshot (first 22+ years of CCA model run):

```{r aggwtcomp}

wtage_annsurv <- catch_length_census_ss$muweight %>%
  filter(time %in% annualmidyear)

# reverse to show agecl time series of wt
wageplot <- ggplot(wtage_annsurv, aes(time, atoutput)) +
  geom_line(aes(colour = factor(agecl))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  xlab("model timestep (5 per year)") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " annual mid year sample"))

wageplot + facet_wrap(c("species"), scales="free_y")
